Concurrent Programming in Harmony

### 2021-02-23 book/paper.tex
Latest

Harmony can report the following failure types:
\begin{itemize}
\item \texttt{Safety violation}: This means something went wrong with
at least one of the executions of the program that it tried.  This
can include a failing assertion, divide by zero, using an uninitialized
or non-existent variable, dividing a set by an integer, and so on.
% It also includes certain infinite loops.
Harmony will print a trace of the shortest bad execution that it found.
\item \texttt{Non-terminating State}: Harmony found one or more states
from which there does not exist an execution such that all threads
terminate.  Harmony will not only print the non-terminating state with
the shortest trace, but also the list of threads
at that state, along with their program counters.
\item \texttt{Active Busy Waiting}:  There are states in which
some thread cannot make progress without the help of another thread but
does not block (see \autoref{ch:condwait});
\item \texttt{Data Race}: There are states in which two or more threads
concurrently access a shared variable, at least one of which is a store
operation (see \autoref{ch:synch}).
% \item \texttt{Stopped States}: Similar to non-terminating states,
% these are states in which some threads are left that cannot
% terminate.  \autoref{ch:synch} will explain what it means for
% a thread to be stopped.
\end{itemize}

Harmony checks for these types of failure conditions in the given
order: if there are multiple failure conditions, only the first is
reported.

In our case, Harmony reports a safety violation.
In particular, it reports that the assertion failed
because \textit{count} has value 1 instead of 2.
Next it reports an execution that failed this assertion.
The program got to the failed assertion in 4 ``turns.''
The output has four columns:
\begin{enumerate}
\item A thread identifier;
\item The main method and argument of the thread;
\item The sequence of program counters of the HVM instructions that the thread executed;
\item The contents of the shared memory.
\end{enumerate}

>>>>

Harmony can report the following failure types:
\begin{itemize}
\item \texttt{Safety violation}: This means something went wrong with
at least one of the executions of the program that it tried.  This
can include a failing assertion, divide by zero, using an uninitialized
or non-existent variable, dividing a set by an integer, and so on.
% It also includes certain infinite loops.
Harmony will print a trace of the shortest bad execution that it found.
\item \texttt{Non-terminating State}: Harmony found one or more states
from which there does not exist an execution such that all threads
terminate.  Harmony will not only print the non-terminating state with
the shortest trace, but also the list of threads
at that state, along with their program counters.
\item \texttt{Active Busy Waiting}:  There are states in which
some thread cannot make progress without the help of another thread but
does not block;
\item \texttt{Data Race}: There are states in which two or more threads
concurrently access a shared variable, at least one of which is a store
operation.
% \item \texttt{Stopped States}: Similar to non-terminating states,
% these are states in which some threads are left that cannot
% terminate.  \autoref{ch:synch} will explain what it means for
% a thread to be stopped.
\end{itemize}

Harmony checks for these types of failure conditions in the given
order: if there are multiple failure conditions, only the first is
reported.  \emph{Active busy waiting} (\autoref{ch:condwait}) is not
technically an indication of a synchronization problem, but instead an
indication of an inefficient solution to a synchronization problem---
one that uses up significantly CPU cycles.
A \emph{data race} (\autoref{ch:synch}) may not be a bug either---whether or
not it is might depend on the semantics of the underlying memory
operations and are therefore generally undesirable.

Going back to our example, Harmony reports a safety violation.
In particular, it reports that the assertion failed
because \textit{count} has value 1 instead of 2.
Next it reports an execution that failed this assertion.
The program got to the failed assertion in 4 ``turns.''
The output has four columns:
\begin{enumerate}
\item A thread identifier;
\item The main method and argument of the thread;
\item The sequence of program counters of the HVM instructions that the thread executed;
\item The contents of the shared memory.
\end{enumerate}
